Machine-learned embedded atom method

### 2019-10-22 doc/papers/ml-eam/manuscript.tex
The manuscript is updated. References are also added.

\author{Xin Chen, 
        Li-Fang Wang, 
        De-Ye Lin$^\mathrm{*}$, 
        Hai-Feng Song$^\mathrm{*}$}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
\end{abstract}

@@

Atomistic modeling plays a vital role in materials science. \textit{ab initio} 
calculation or force-field based molecular dynamics simulation (MD) are 
effective ways to study, understand or predict chemical and physical properties 
of materials. \textit{ab initio} approaches are generally much more precise but 
they are rarely used on large-scale metallic systems due to their extremely-high 
computation expenses. Physical model based empirical potential (force-field), 
such as the embedded-atom method (EAM), modified embedded-atom method (MEAM), 
bond-order potential (BoP), or angular-dependent potential (ADP), still plays 
the major role in long-time simulations and these empirical methods can achieve 
reasonable accuracy with much lower computation costs. Empirical potentials 
generally have very few learnable parameters and both microscopic observables 
(energy, forces, virial, etc.) and macroscopic observables (melting 
point, surface energy, etc.) can be used to tune these parameters. Finding 
optimal parameters of empirical potentials is always a challenging task. Global 
optimization (GO) approaches (Basin-Hopping, genetic algorithm, etc) are 
traditionally used to find the best possible parameters. However, the gradients 
of the losses with respect to model parameters are difficult or even impossible 
to calcualte. Hence, GO optimizations are generally not that effective.

In the last decade, machine learning (ML) has become one of the hottest 
topics in many research areas. In the materials science, researchers have made 
great efforts on developing ML models to describe atomic interactions. Such ML 
models are considered as machine learning interaction potentials (MLIPs). Until 
now, hundreds of MLIPs have been proposed. Among them, the symmetry-function 
based atomistic neural network (ANN) model, published by Parinello and Behler in 
2007, is still the most popular choice in modeling metallic interactions. 
The smooth overlap atomic positions descriptor based gaussian approximation 
potential (SOAP-GAP), developed by Bartòk et al, can give extremely accurate 
prediction results, although it's a bit computationally expensive. 
Recently, Thompson and co-workers proposed another quantum-accurate MLIP named 
the spectral neighbor analysis potential (SNAP) and it has been proven working  
on a broad range of metals and alloys.

In many cases, MILPs can easily outperform state-of-art empirical potentials. 
Compared with empirical potentials, MLIPs generally have orders of magnitudes 
more model paramters. The redundant parameter space greatly reduces the 
difficulty of fitting complicated potential energy surfaces. But, to effectively 
train a MLIP and avoid overfitting, a large high-quality (versatile) training 
dataset is probably needed. However, MILPs \textbf{can really} take advantages 
of "big data" for two reasons. First, MLIPs typically only have 
\textit{basic} or \textit{simple} arithmetic operations. Thus, MILPs can be 
implemented within modern deep learning frameworks (TensorFlow, PyTorch, etc) 
so that the gradients of the total loss with respect to fitting parameters can 
be can obtained with the backpropagation algorithm automatically and 
efficiently. Second, MLIPs are mostly vectorizable. Hence, GPUs can be utilized
to significantly accelerate training and using of MILPs.

However, MLIPs also have challenges. The large parameter space and lack of 
physical background makes the "big data" a necessarity. The cost of dataset is 
non-negligible. Besides, even "big data" can only cover a small portion of real 
physical environments (temperature, external pressure, etc). Outside the 
training zone, the performances of MILPs may not that stable. For long-time 
molecular dynamics (MD) simulations of large-scale ($10^5$ or more) systems, 
computation efficiency also becomes a major concern. Recent benchmark tests 
suggest that MILPs are still too expensive. At present, most MLIPs are used to 
examine small to medium ($10^3$ to $10^4$) systems.

@@

This paper is organized as follows. Section \ref{sec:method} describes the 
theoretical background of this work, including the formalism of the embedded 
atom method and algorithms and details of the machine learned EAM. 
Section \ref{sec:results} summarizes the training results and the optimal 
parameters. Applications of the new potentials are discussed in 
Section \ref{sec:discussions}.

@@

In the original EAM formalism, the total energy, $E^{total}$, is the sum of 
atomic energies:
\begin{align}
E^{total} = & \sum_{i}^{N}{E_{i}} \nonumber \\
\label{eq:eam_total_energy}
= & \sum_{i}^{N}{F_{a}(\rho_i)} + 
    \frac{1}{2}\sum_{i}^{N}{\sum_{j \neq i}^{r_{ij} < r_c}{
    \phi_{ab}(r_{ij})
}}
\end{align}
where $r_c$ is the cutoff radius, $a$ and $b$ represents species of atoms $i$ 
and $j$, $\phi_{ab}(r_{ij})$ is energy of the pairwise interaction between $i$ 
and $j$, $F_{a}(\rho_{i})$ is the embedding energy and $\rho_{i}$ is the local 
electron density of atom $i$. $\rho_{i}$ can be calculated with the following 
equation:
\begin{equation}
\label{eq:rho_eam}
\rho_{i} = \sum_{j}^{r_{ij} < r_{c}}{
    \rho_{b}(r_{ij})
}
\end{equation}
where $\rho_{b}$ is the electron density function of specie $b$. In the 
Finnis-Sinclair model, the electron density has a slightly modified form:
\begin{equation}
\label{eq:rho_fs}
\rho_{i}^{\mathrm{FS}} = \sum_{j}^{r_{ij} < r_{c}}{
    \rho_{ab}(r_{ij})
}
\end{equation}
$F$, $\rho$ and $\phi$ can be either parameterized functions or cubic splines.

The original EAM formalism does not inlcude angular-dependent interactions. To 
fix this problem, Baskes modified the original EAM and got MEAM (modified 
embedded-atom method), Lenosky proposed an alternative spline-based interpretion 
of MEAM  while Mishin developed the angular-dependent potential (ADP). The ADP 
formalism introduces three additional terms to the total energy:
\begin{align}
\label{eq:adp}
E^{total} 
& = E^{\mathrm{EAM}} \nonumber \\
& + \frac{1}{2}\sum_{i}{\sum_{\alpha}{(\mu_{i}^{\alpha})^2}} \nonumber \\
& + \frac{1}{2}\sum_{i}{
    \sum_{\alpha}{\sum_{\beta}{(\lambda_{i}^{\alpha\beta})^2}}} \nonumber \\
& - \frac{1}{6}\sum_{i}{\nu_{i}^{2}}
\end{align}
These terms represent non-central bonding contributions and they can be computed
with the following equations:
\begin{align}
\label{eq:adp_mu}
\mu_{i}^{\alpha} & = \sum_{j \neq i}{\mu_{ab}(r_{ij}) r_{ij}^{\alpha}} \\
\label{eq:adp_lambda}
\lambda_{i}^{\alpha\beta} & = \sum_{j \neq i}{
    \omega_{ab}(r_{ij}) r_{ij}^{\alpha}r_{ij}^{\beta}} \\
\label{eq:adp_nu}
\nu_{i} & = \sum_{\alpha}{\lambda_{i}^{\alpha\alpha}}
\end{align}
where $\mu_{ab}(r)$ and $\omega_{ab}(r)$ can be viewed as measures of the 
strengths of dipole and quadrupole interactions.

@@

In this work, we use the EAM potential published by Zhou, Johnson and Wadley 
(Zjw04) as an example to validate our machine learning approach. Zjw04 is a 
quite popular EAM potential. In the Zjw04 potential, the electron density 
function has the following form:
\begin{equation}
\label{eq:zjw04_rho}
\rho_{b}(r) = \frac{
    f_{e} \exp\left[ -\beta\left( r/r_{e} - 1 \right) \right]
}{
    1 + \left(r/r_{e} - \lambda\right)^{20}
}
\end{equation}
where $r_e$ is a constant equal to equilibrium spacing between nearest neighors, 
$f_{e}$, $\beta$, $\lambda$ are adjustable parameters. The pairwise potential 
between the same species can be computed with:
\begin{equation}
\label{eq:zjw04_phi_aa}
\phi_{aa}(r) = 
\frac{A \exp\left[ -\alpha\left( r/r_{e} - 1 \right) \right]}
     {1 + \left(r / r_{e} - \kappa\right)^{20}} - 
\frac{B \exp\left[ -\beta\left( r/r_{e} - 1 \right) \right]}
     {1 + \left(r / r_{e} - \lambda\right)^{20}}
\end{equation}
where $A$, $B$, $\alpha$ and $\kappa$ are also trainable parameters, $\beta$ and 
$\kappa$ are used in Equation \ref{eq:zjw04_rho} already. For the pairwise 
interaction between two atoms of different species, Zhou et al proposed an 
interpolation form:
\begin{equation}
\label{eq:zjw04_phi_ab}
\phi_{ab}(r) = \frac{1}{2}\left(
    \frac{\rho_{b}(r)}{\rho_{a}(r)}\phi_{aa}(r) +
    \frac{\rho_{a}(r)}{\rho_{b}(r)}\phi_{bb}(r) 
\right)
\end{equation}
The embedding function has a more complicated form as it requires to fit a much 
wider range of electron density values:
\begin{equation}
\label{eq:zjw04_embed}
F(\rho) = \begin{cases}
    \sum_{i=0}^{3}{F_{ni}\left( \frac{\rho}{\rho_n} - 1 \right)^{i}} &
    \rho < \rho_{n} \\
    \sum_{i=0}^{3}{F_{i}\left( \frac{\rho}{\rho_e} - 1 \right)^{i}} &
    \rho_{n} \leq \rho < \rho_{0} \\
    F_{e}\left[ 
        1 - \eta\ln\left( \frac{\rho}{\rho_s}\right) 
    \right](\frac{\rho}{\rho_s})^{\eta} & \rho_{0} \leq \rho
\end{cases}
\end{equation}
where $F_{ni}$, $F_{i}$, $\rho_e$, $\rho_s$, $\eta$ and $F_{e}$ are trainable 
parameters, $\rho_n=0.85\rho_e$ and $\rho_0=1.15\rho_e$. For each metal, there 
are 15 adjustable parameters. The original embedding potential is a stepwise 
function. Thus, the minimization requires some tricks to ensure its continuity.
To make it simpler, we slightly modified Equation \ref{eq:zjw04_embed}:
\begin{align}
\label{eq:zjw04xc_embed}
F(\rho) 
= & c_1 \cdot
\sum_{i=0}^{3}{F_{ni}\left( \frac{\rho}{\rho_n} - 1 \right)^{i}} \nonumber \\
+ & c_2 \cdot 
\sum_{i=0}^{3}{F_{i}\left( \frac{\rho}{\rho_e} - 1 \right)^{i}} \nonumber \\
+ & c_3 \cdot
F_{e}\left[1 - \eta\ln\left( \frac{\rho}{\rho_s}\right)\right]
(\frac{\rho}{\rho_s})^{\eta} \\
\label{eq:frho_c1}
c_{1} = & \frac{1}{1 + e^{-2\left(\rho_{n} - \rho\right)}} \\
\label{eq:frho_c3}
c_{3} = & \frac{1}{1 + e^{-2\left(\rho - \rho_{0}\right)}} \\
c_{2} = & 1 - c_1 - c_3
\end{align}
$\omega_1$ and $\omega_3$ are just damping factors calculated by the sigmoid 
functions (Equations \ref{eq:frho_c1} and \ref{eq:frho_c3}).

The dipole ($\mu_{ab}$) and quadrupole ($\lambda_{ab}$) functions have the same 
form (developed by Mishin):
\begin{align}
\label{eq:adp_dipole}
\mu_{ab}(r) & = \left[
    d_{1}^{ab} \exp\left( -d_{2}^{ab}r \right) + d_{3}^{ab}
\right] \psi\left( \frac{r - r_{0}}{r_{h}} \right) \\
\omega_{ab}(r) & = \left[
    q_{1}^{ab} \exp\left( -q_{2}^{ab}r \right) + q_{3}^{ab}
\right] \psi\left( \frac{r - r_{0}}{r_{h}} \right)
\end{align}
where $d_{i}$, $q_{i}$, $r_0$ and $r_{h}$ are trainable parameters and $\psi(x)$
is a damping function:
\begin{equation}
\label{eq:mishin_cutoff}
\psi(x) = \begin{cases}
    0 & x \ge 0 \\
    \frac{x^4}{1 + x^4} & x < 0 
\end{cases}
\end{equation}

@@

In this work, we successfully integrated two constraints into the total loss:
the Rose equation of state (EOS) constraint and the elastic tensor constraint. 
These two losses will be discussed later. The details of their implementations 
will be described in another paper.

@@

In this work, we successfully find a way to use elastic tensor directly as a 
constraint. Given an equilibrium crystal structure, its elastic constant 
$c_{ijkl}$ can be derived from $E^{total}$ directly:
\begin{align}
\label{eq:virial}
V \cdot \epsilon & = -\mathbf{F}^{T}\mathbf{R} + \left(
    \frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h} \\ 
\label{eq:cijkl}
c_{ijkl} |_{\epsilon \to 0, \mathbf{F} \to 0} & = \frac{1}{V}\left[
    \left( 
        \frac{\partial{\epsilon_{ij}}}{\partial{\mathbf{h}}}
    \right)^{\mathrm{T}}\mathbf{h}
\right]_{kl}
\end{align}
where $V$ is the volume, $\epsilon$ is the $3 \times 3$ virial stress tensor, 
$\mathbf{h}$ is the \textit{row-major} $3 \times 3$ lattice tensor, $\mathbf{F}$ 
and $\mathbf{R}$ are $N \times 3$ matrices representing the total forces and 
atomic positions.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}
\centering
\begin{tabular}{lccccc}
\hline
                              & DFT  & SNAP          & EAM            & ML-EAM         & Exp.      \\
\hline
$T_{m}$(K)                    &      & 1785          & 1520           & 1520           & 1728      \\
$c_{11}$ (GPa)                & 276  & 276 (0.0\%)   & 248 (-10.1\%)  & 274 (-0.7\%)   & 261       \\
$c_{12}$ (GPa)                & 159  & 159 (0.0\%)   & 147 (-7.5\%)   & 163 (2.5\%)    & 151       \\
$c_{44}$ (GPa)                & 132  & 132 (0.0\%)   & 125 (-5.3\%)   & 131 (-0.8\%)   & 132       \\
$B_{\mathrm{VRH}}$ (GPa)      & 198  & 198 (0.0\%)   & 181 (-8.6\%)   & 195 (0.0\%)    & 188       \\
$G_{\mathrm{VRH}}$ (GPa)      & 95   & 95 (0.0\%)    & 87 (-8.4\%)    & 93 (-2.1\%)    & 93        \\
$\mu$                         & 0.29 & 0.29 (0.0\%)  & 0.29 (0.0\%)   & 0.29 (0.0\%)   & 0.29      \\
$E_{v}$ (eV)                  & 1.46 & 1.68 (15.1\%) & 1.68 (15.1\%)  & 1.71 (17.1\%)  & 1.54–1.80 \\
$E_{m}$ (eV)                  & 1.12 & 1.07 (-4.5\%) & 0.90 (-19.6\%) & 0.87 (-22.3\%) & 1.01–1.48 \\
$E_{a} = E_{v} + E_{m} $ (eV) & 2.58 & 2.75 (6.6\%)  & 2.58 (0.0\%)   & 2.58 (0.0\%)   & 2.77–2.95 \\
\hline
\end{tabular}
\caption{\label{table:Ni_properties} Comparison of the calculated and 
experimental material properties of elementary fcc Ni. $T_{m}$ is the melting 
point. $c_{ij}$ represents elastic constants. $E_{v}$ and $E_{m}$ are vacancy 
formation energy and migration energy, respectively. SNAP refers to the 
elementary Ni SNAP model.
}
\end{table*}

@@

The publicly Ni-Mo dataset is used to evaluate our approach. This dataset is 
provided by Shyue Ping Ong. It contains 3973 different Ni-Mo solids. All 
calculations were done by VASP with the PBE functional and the projector 
augmented-wave approach.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: r(Mo-Mo)
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.64]{figures/r_MoMo_hist.png}
\caption{\label{fig:r_MoMo_hist} The energy-volume curves of the original EAM, 
SNAP, ML-EAM and DFT for the fcc Ni.}
\end{figure}

>>>>

\author{Xin Chen, 
        Li-Fang Wang, 
        Xing-Yu Gao,
        Ya-Fan Zhao,
        De-Ye Lin$^\mathrm{*}$, 
        Wei-Dong Chu,
        Hai-Feng Song$^\mathrm{*}$}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Embedded atom method (EAM) and its variant angular-dependent potential (ADP) 
have proven successful in modeling fcc and bcc metals qualitatively. In the past 
few years, machine learning has became the central approach for modeling 
interatomic interactions. Powered by big data and efficient optimizers, machine 
learning interatomic potentials can generally approximate first-principle 
accuracy. In this works, we successfully integrated machine learning with EAM 
and ADP. Machine learning can significantly boost their performances. 
For energy and force predictions, machine leaned EAM and ADP can be almost 
as accurate as the spectral neighbor analysis potential (SNAP) on the fcc Ni, 
bcc Mo and Mo-Ni alloy systems, while the later approach is nealy 1000 times 
slower. ML-EAM and ML-ADP can also give first-principle accuracy key material 
properties, such as elastic constants, melting temperatures. Our results suggest 
a new and systematic route for developing machine learning interatomic 
potentials. 
Our 
\end{abstract}

@@

Atomistic modeling plays a vital role in materials science. \textit{ab initio} 
calculation or force-field based molecular dynamics simulation (MD) are 
effective ways to study, understand or predict chemical and physical properties 
of materials. \textit{ab initio} approaches are generally much more precise but 
they are rarely used on large-scale metallic systems due to their extremely-high 
computation expenses. Physical model based empirical potential (force-field), 
such as the embedded-atom method (EAM) \cite{EAM0,EAM1,EAM2,EAM3,EAM4}, 
modified embedded-atom method (MEAM) \cite{MEAM0,MEAM1,MEAM2,MEAM_Jelinek}, 
bond-order potential (BoP) \cite{BOP0,BOP1,BOP2,BOP3}, or angular-dependent 
potential (ADP) \cite{ADP0,ADP1,ADP2,ADP3,ADP_FeNi}, still plays the major role 
in long-time simulations and these empirical methods can achieve reasonable 
accuracy with much lower computation costs. Empirical potentials 
generally have very few learnable parameters and both microscopic observables 
(energy, forces, virial, etc.) and macroscopic observables (melting 
point, surface energy, etc.) can be used to tune these parameters. Finding 
optimal parameters of empirical potentials is always a challenging task. Global 
optimization (GO) approaches (Basin-Hopping \cite{BH0,BH1}, genetic algorithm 
\cite{PCIPY,PGA}, etc) are traditionally used to find the best possible 
parameters. However, the gradients of the losses with respect to model 
parameters are difficult or even impossible to calcualte. Hence, 
GO optimizations are generally not that effective.

In the last decade, machine learning (ML) has become one of the hottest 
topics in many research areas. In the materials science, researchers have made 
great efforts on developing ML models to describe atomic interactions. Such ML 
models are considered as machine learning interaction potentials (MLIPs). Until 
now, hundreds of MLIPs have been proposed. Among them, the symmetry-function 
based atomistic neural network (ANN) model, published by Parinello and Behler in 
2007 \cite{SF2007,SF1,SF2,SF3,SF4}, is still the most popular choice in modeling 
metallic interactions \cite{SFApp0,SFApp1,SFApp2}. 
The smooth overlap atomic positions descriptor based gaussian approximation 
potential (SOAP-GAP) \cite{soap0,soap1,soap2,soap3}, developed by Bartòk et al, 
can give extremely accurate prediction results, although it's a bit 
computationally expensive. 
Recently, Thompson and co-workers proposed another quantum-accurate MLIP named 
the spectral neighbor analysis potential (SNAP) \cite{SNAP_Algo} and it has 
been proven working on a broad range of metals and 
alloys \cite{SNAP_Mo,SNAP,SNAP_New}.

In many cases, MILPs can easily outperform state-of-art empirical potentials. 
Compared with empirical potentials, MLIPs generally have orders of magnitudes 
more model paramters. The redundant parameter space greatly reduces the 
difficulty of fitting complicated potential energy surfaces. But, to effectively 
train a MLIP and avoid overfitting, a large high-quality (versatile) training 
dataset is probably needed. However, MILPs \textbf{can really} take advantages 
of "big data" for two reasons. First, MLIPs typically only have 
\textit{basic} or \textit{simple} arithmetic operations. Thus, MILPs can be 
implemented within modern deep learning frameworks 
(TensorFlow \cite{tensorflow}, PyTorch \cite{pytorch}, etc) 
so that the gradients of the total loss with respect to fitting parameters can 
be can obtained with the backpropagation algorithm automatically and 
efficiently. Second, MLIPs are mostly vectorizable. Hence, GPUs can be utilized
to significantly accelerate training and using of MILPs.

However, MLIPs also have challenges. The large parameter space and lack of 
physical background makes the "big data" a necessarity. The cost of dataset is 
non-negligible. Besides, even "big data" can only cover a small portion of real 
physical environments (temperature, external pressure, etc). Outside the 
training zone, the performances of MILPs may not that stable. For long-time 
molecular dynamics (MD) simulations of large-scale ($10^5$ or more) systems, 
computation efficiency also becomes a major concern. Recent benchmark 
tests \cite{SnapBenchmark} suggest that MILPs are still too expensive. 
At present, most MLIPs are used to examine small to medium ($10^3$ to $10^4$) 
systems \cite{SNAP,SFApp0,SFApp1,SFApp2,soap3}.

@@

This paper is organized as follows. Section \ref{sec:method} describes the 
theoretical background of this work, including the formalism of the embedded 
atom method and algorithms of the machine learned EAM. Section \ref{sec:results} 
summarizes the results of ML-EAM and ML-ADP. Further discussions are given in 
section \ref{sec:discussions}.

@@

In the original EAM formalism \cite{EAM0}, the total energy, $E^{total}$, is the 
sum of atomic energies:
\begin{align}
E^{total} = & \sum_{i}^{N}{E_{i}} \nonumber \\
\label{eq:eam_total_energy}
= & \sum_{i}^{N}{F_{a}(\rho_i)} + 
    \frac{1}{2}\sum_{i}^{N}{\sum_{j \neq i}^{r_{ij} < r_c}{
    \phi_{ab}(r_{ij})
}}
\end{align}
where $r_c$ is the cutoff radius, $a$ and $b$ represents species of atoms $i$ 
and $j$, $\phi_{ab}(r_{ij})$ is energy of the pairwise interaction between $i$ 
and $j$, $F_{a}(\rho_{i})$ is the embedding energy and $\rho_{i}$ is the local 
electron density of atom $i$. $\rho_{i}$ can be calculated with the following 
equation:
\begin{equation}
\label{eq:rho_eam}
\rho_{i} = \sum_{j}^{r_{ij} < r_{c}}{
    \rho_{b}(r_{ij})
}
\end{equation}
where $\rho_{b}$ is the electron density function of specie $b$. 
$F$, $\rho$ and $\phi$ can be either parameterized functions or cubic splines.

The original EAM does not inlcude angular-dependent interactions. To 
fix this problem, Baskes modified the original EAM and got MEAM (modified 
embedded-atom method) \cite{MEAM0}, Lenosky proposed an alternative spline-based 
interpretion of MEAM \cite{MEAM_Lenosky} while Mishin developed the 
angular-dependent potential (ADP) \cite{ADP0,ADP1,ADP2,ADP3,ADP4}. The ADP 
formalism introduces three additional terms to the total energy:
\begin{align}
\label{eq:adp}
E^{total} 
& = E^{\mathrm{EAM}} \nonumber \\
& + \frac{1}{2}\sum_{i}{\sum_{\alpha}{(\mu_{i}^{\alpha})^2}} \nonumber \\
& + \frac{1}{2}\sum_{i}{
    \sum_{\alpha}{\sum_{\beta}{(\lambda_{i}^{\alpha\beta})^2}}} \nonumber \\
& - \frac{1}{6}\sum_{i}{\nu_{i}^{2}}
\end{align}
These terms represent non-central bonding contributions and they can be computed
with the following equations:
\begin{align}
\label{eq:adp_mu}
\mu_{i}^{\alpha} & = \sum_{j \neq i}{\mu_{ab}(r_{ij}) r_{ij}^{\alpha}} \\
\label{eq:adp_lambda}
\lambda_{i}^{\alpha\beta} & = \sum_{j \neq i}{
    \omega_{ab}(r_{ij}) r_{ij}^{\alpha}r_{ij}^{\beta}} \\
\label{eq:adp_nu}
\nu_{i} & = \sum_{\alpha}{\lambda_{i}^{\alpha\alpha}}
\end{align}
where $\mu_{ab}(r)$ and $\omega_{ab}(r)$ can be viewed as measures of the 
strengths of dipole and quadrupole interactions.

@@

In this work, we use the EAM potential published by Zhou, Johnson and Wadley 
(Zjw04) \cite{ZJW0,ZJW1,ZJW2} as an example to validate our machine learning 
approach. Zjw04 is a quite popular EAM potential. In the Zjw04 potential, the 
electron density function has the following form:
\begin{equation}
\label{eq:zjw04_rho}
\rho_{b}(r) = \frac{
    f_{e} \exp\left[ -\beta\left( r/r_{e} - 1 \right) \right]
}{
    1 + \left(r/r_{e} - \lambda\right)^{20}
}
\end{equation}
where $r_e$ is a constant equal to equilibrium spacing between nearest neighors, 
$f_{e}$, $\beta$, $\lambda$ are adjustable parameters. The pairwise potential 
between the same species can be computed with:
\begin{equation}
\label{eq:zjw04_phi_aa}
\phi_{aa}(r) = 
\frac{A \exp\left[ -\alpha\left( r/r_{e} - 1 \right) \right]}
     {1 + \left(r / r_{e} - \kappa\right)^{20}} - 
\frac{B \exp\left[ -\beta\left( r/r_{e} - 1 \right) \right]}
     {1 + \left(r / r_{e} - \lambda\right)^{20}}
\end{equation}
where $A$, $B$, $\alpha$ and $\kappa$ are also trainable parameters, $\beta$ and 
$\kappa$ are used in Equation \ref{eq:zjw04_rho} already. For the pairwise 
interaction between two atoms of different species, Zhou et al proposed an 
interpolation form:
\begin{equation}
\label{eq:zjw04_phi_ab}
\phi_{ab}(r) = \frac{1}{2}\left(
    \frac{\rho_{b}(r)}{\rho_{a}(r)}\phi_{aa}(r) +
    \frac{\rho_{a}(r)}{\rho_{b}(r)}\phi_{bb}(r) 
\right)
\end{equation}
The embedding function has a more complicated form as it requires to fit a much 
wider range of electron density values:
\begin{equation}
\label{eq:zjw04_embed}
F(\rho) = \begin{cases}
    \sum_{i=0}^{3}{F_{ni}\left( \frac{\rho}{\rho_n} - 1 \right)^{i}} &
    \rho < \rho_{n} \\
    \sum_{i=0}^{3}{F_{i}\left( \frac{\rho}{\rho_e} - 1 \right)^{i}} &
    \rho_{n} \leq \rho < \rho_{0} \\
    F_{e}\left[ 
        1 - \eta\ln\left( \frac{\rho}{\rho_s}\right) 
    \right](\frac{\rho}{\rho_s})^{\eta} & \rho_{0} \leq \rho
\end{cases}
\end{equation}
where $F_{ni}$, $F_{i}$, $\rho_e$, $\rho_s$, $\eta$ and $F_{e}$ are trainable 
parameters, $\rho_n=0.85\rho_e$ and $\rho_0=1.15\rho_e$. For each metal, there 
are 15 adjustable parameters. The original embedding potential is a stepwise 
function. Thus, the minimization requires some tricks to ensure its continuity.
To make it simpler, we slightly modified Equation \ref{eq:zjw04_embed}:
\begin{align}
\label{eq:zjw04xc_embed}
F(\rho) 
= & c_1 \cdot
\sum_{i=0}^{3}{F_{ni}\left( \frac{\rho}{\rho_n} - 1 \right)^{i}} \nonumber \\
+ & c_2 \cdot 
\sum_{i=0}^{3}{F_{i}\left( \frac{\rho}{\rho_e} - 1 \right)^{i}} \nonumber \\
+ & c_3 \cdot
F_{e}\left[1 - \eta\ln\left( \frac{\rho}{\rho_s}\right)\right]
(\frac{\rho}{\rho_s})^{\eta} \\
\label{eq:frho_c1}
c_{1} = & \frac{1}{1 + e^{-2\left(\rho_{n} - \rho\right)}} \\
\label{eq:frho_c3}
c_{3} = & \frac{1}{1 + e^{-2\left(\rho - \rho_{0}\right)}} \\
c_{2} = & 1 - c_1 - c_3
\end{align}
$\omega_1$ and $\omega_3$ are just damping factors calculated by the sigmoid 
functions (Equations \ref{eq:frho_c1} and \ref{eq:frho_c3}).

The dipole ($\mu_{ab}$) and quadrupole ($\lambda_{ab}$) functions have the same 
form (developed by Mishin \cite{ADP0}):
\begin{align}
\label{eq:adp_dipole}
\mu_{ab}(r) & = \left[
    d_{1}^{ab} \exp\left( -d_{2}^{ab}r \right) + d_{3}^{ab}
\right] \psi\left( \frac{r - r_{0}}{r_{h}} \right) \\
\omega_{ab}(r) & = \left[
    q_{1}^{ab} \exp\left( -q_{2}^{ab}r \right) + q_{3}^{ab}
\right] \psi\left( \frac{r - r_{0}}{r_{h}} \right)
\end{align}
where $d_{i}$, $q_{i}$, $r_0$ and $r_{h}$ are trainable parameters and $\psi(x)$
is a damping function:
\begin{equation}
\label{eq:mishin_cutoff}
\psi(x) = \begin{cases}
    0 & x \ge 0 \\
    \frac{x^4}{1 + x^4} & x < 0 
\end{cases}
\end{equation}

@@

In this work, we successfully integrated two constraints into the total loss:
the Rose equation of state (EOS) \cite{Rose0,Rose1,Rose2} constraint and the 
elastic tensor constraint. These two losses will be discussed later. The details 
of their implementations will be described in another paper.

@@

In this work, we successfully find a way to use elastic tensor directly as a 
constraint. Given an equilibrium crystal structure, its elastic constant 
$c_{ijkl}$ can be derived from $E^{total}$ directly \cite{Elastic}:
\begin{align}
\label{eq:virial}
V \cdot \epsilon & = -\mathbf{F}^{T}\mathbf{R} + \left(
    \frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h} \\ 
\label{eq:cijkl}
c_{ijkl} |_{\epsilon \to 0, \mathbf{F} \to 0} & = \frac{1}{V}\left[
    \left( 
        \frac{\partial{\epsilon_{ij}}}{\partial{\mathbf{h}}}
    \right)^{\mathrm{T}}\mathbf{h}
\right]_{kl}
\end{align}
where $V$ is the volume, $\epsilon$ is the $3 \times 3$ virial stress tensor, 
$\mathbf{h}$ is the \textit{row-major} $3 \times 3$ lattice tensor, $\mathbf{F}$ 
and $\mathbf{R}$ are $N \times 3$ matrices representing the total forces and 
atomic positions.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}
\centering
\begin{tabular}{lccccc}
\hline
                              & DFT  & SNAP \cite{SNAP} & EAM              & ML-EAM         & Exp.                           \\
\hline
$T_{m}$(K)                    &      & 1785             & 1520 \cite{ZJW2} & 1520           & 1728                           \\
$c_{11}$ (GPa)                & 276  & 276 (0.0\%)      & 248 (-10.1\%)    & 274 (-0.7\%)   & 261 \cite{Ni_Elastic_Exp}      \\
$c_{12}$ (GPa)                & 159  & 159 (0.0\%)      & 147 (-7.5\%)     & 163 (2.5\%)    & 151 \cite{Ni_Elastic_Exp}      \\
$c_{44}$ (GPa)                & 132  & 132 (0.0\%)      & 125 (-5.3\%)     & 131 (-0.8\%)   & 132 \cite{Ni_Elastic_Exp}      \\
$B_{\mathrm{VRH}}$ (GPa)      & 198  & 198 (0.0\%)      & 181 (-8.6\%)     & 195 (0.0\%)    & 188                            \\
$G_{\mathrm{VRH}}$ (GPa)      & 95   & 95 (0.0\%)       & 87 (-8.4\%)      & 93 (-2.1\%)    & 93                             \\
$\mu$                         & 0.29 & 0.29 (0.0\%)     & 0.29 (0.0\%)     & 0.29 (0.0\%)   & 0.29                           \\
$E_{v}$ (eV)                  & 1.46 & 1.68 (15.1\%)    & 1.68 (15.1\%)    & 1.71 (17.1\%)  & 1.54–1.80 \cite{Ni_EvEmEa_Exp} \\
$E_{m}$ (eV)                  & 1.12 & 1.07 (-4.5\%)    & 0.90 (-19.6\%)   & 0.87 (-22.3\%) & 1.01–1.48 \cite{Ni_EvEmEa_Exp} \\
$E_{a} = E_{v} + E_{m} $ (eV) & 2.58 & 2.75 (6.6\%)     & 2.58 (0.0\%)     & 2.58 (0.0\%)   & 2.77–2.95 \cite{Ni_EvEmEa_Exp} \\
\hline
\end{tabular}
\caption{\label{table:Ni_properties} Comparison of the calculated and 
experimental material properties of elementary fcc Ni. $T_{m}$ is the melting 
point. $c_{ij}$ represents elastic constants. $E_{v}$ and $E_{m}$ are vacancy 
formation energy and migration energy, respectively. SNAP refers to the 
elementary Ni SNAP model.
}
\end{table*}

@@

The publicly Ni-Mo dataset \cite{SNAP} is used to evaluate our approach. This 
dataset is provided by Shyue Ping Ong. It contains 3973 different Ni-Mo solids. 
All calculations were done by VASP \cite{VASP} with the PBE \cite{PBE} 
functional and the projector augmented-wave approach \cite{PAW}.

@@

Figure \ref{fig:Mo_eam_adp} compares the $\rho(r)$, $F(\rho)$, $\phi(r)$, 
$\mu(r)$ and $\omega(r)$ functions. The differences between ML-EAM and ML-ADP 
$\rho(r)$, $\phi(r)$ and $F(\rho)$ are quite small. Both the angular-dependent 
dipole $\mu(r)$ and quadrupole $\omega(r)$ quickly descrease to zero at 2.5 \AA, 
which is lower than the equilibrium spacing (2.7 \AA) between two adjacent Mo 
atoms. Hence, the dipole and quadrupole terms only contribute a little to the 
total energy for this test case.
